---
title: "Requests on the web service 2 - OpenSILEX"
author: "I.Sanchez, A. Charleroy, Jean-Eudes Hollebecq (INRA MISTEA)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Requests on the web service 2 - OpenSILEX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!---
Program  : introWS2.Rmd
Author   : I.Sanchez (MISTEA)
Objective: a vignette for phisWSClientR library
           an introduction to requesting the phenomeapi (phis web service) WS2
Creation : 24/01/2019
Update   : 14/02/2019
-->

***

<div style="background-color:rgba(0, 255, 0,0.1); text-align:left; vertical-align: center; padding:10px 0;">
This tutorial explains how to access to the second **phenomeapi** web service (PHIS system information).

For this second web service, the available functions are:

* getAnnotations()
* getEnvironmentData()
* getEvents()
* getExperiments2()
* getInfrastructures()
* getMethods2()
* getPhenotypeData()
* getProjects2()
* getProvenances()
* getScientificObjects()
* getSensors()
* getSpecies()
* getTraits()
* getUnits()
* getVariables2()
* getVectors()
</div>
```{r option}
knitr::opts_chunk$set( echo = TRUE, message = FALSE, warning = FALSE)
```


```{r library, echo=TRUE,message=FALSE, warning=FALSE}
  library(phisWSClientR)
```

# Generalities

## HTTP request
This package performs request to PHIS Web Service. We can distinguish two kinds of request:  
-  GET request that retrieve data from the web service.  
-  POST request that send data to the web service.  

They are nammed accordingly to the HTTP method GET and POST associated to it.

Because these functions rely on HTTP method, the first answer of the functions is the status of the request:  
&nbsp; &nbsp; 200 and family: means the request has succeded  
&nbsp; &nbsp; 300 and family: means the request was reoriented elsewhere  
&nbsp; &nbsp; 400 and family: means the request failed from your side (your fault)  
&nbsp; &nbsp; 500 and family: means that the request failed from the server (not your fault)  
  
What is distinct between the functions is the value taken in the $data list.  
Because this package tries to provide the most generic functions, the format has to be compatible with a lot of data structures. For examples list containing another list (a properties argument typically). That is to say, data.frames are not always compliant enough.

## Identifiers

PHIS relies on URI (Uniform Resource Identifier) to identify resources. As an identifier, this is an un-ambiguous way to identify objects.  

When collecting data through the `getData` function, you will need to retrieve metadata of the variables through the `getVariables2` function if you want the label, unit and method used to create those data.

When inserting new data in the information system, a URI will be automaticaly given to the data/object/resource, based on a certain key (variable+date+provenance for a data value). Some functions will allow you to specify the URI of the resource when creating it. That is the case if the resource has been created somewhere else before.


# Creation of a connection

Ask permission to request to the web service:

```{r connect}
  # If you want to access to a private web service, you have to insert the address of the WS and the port
  #connectToPHISWS(apiID="ws_private",username="guest@opensilex.org",password="guest", url = "www.opensilex.org/openSilexAPI/rest/")
  connectToPHISWS(apiID="ws_private",username="guest@opensilex.org",password="guest", url = "http://www.opensilex.org/openSilexSandBoxAPI/rest/")
```

*aToken* is your identification and grants the access to some data according to your login information.

```{r token}
  aToken <- opensilexWSClientR::getToken("guest@opensilex.org","guest")
```

This token is handled automaticaly for you, but you can acces it using the following command :

```{r get user info}
getUserInformations()
```

# Fetch informations about elements in the web service

## Projects

```{r getProject}
  getProjects2( acronyme = "PHENOME-FPPN")$data
```

## Experiments

```{r getExperiments}
  getExperiments2( uri="http://www.opensilex.org/demo/DIA2017-1")
```
## Scientific Object

```{r getScientificObjects}
  getScientificObjects( uri="http://www.opensilex.org/demo/2018/o18000076")
```

## Infrastructures

```{r getInfrastructures}
  getInfrastructures(uri="https://emphasis.plant-phenotyping.eu")
```

## Variables 

```{r getVariables}
  getVariables2( uri="http://www.opensilex.org/demo/id/variables/v010")
```

## Traits

```{r getTraits}
  getTraits( uri="http://www.opensilex.org/demo/id/traits/t010")
```

## Methods

```{r getMethods}
  getMethods2( uri="http://www.opensilex.org/demo/id/methods/m010")
```

## Units

```{r getUnits}
  getUnits( uri="http://www.opensilex.org/demo/id/units/u007")
```

## Provenances

```{r getProvenances}
getProvenances(label = "ImageAnalysis-ZA17-Maize" )
```


## Species

```{r getSpecies}
  getSpecies( uri="http://www.opensilex.org/id/species/helianthusannuus", language = "en")
```

## Sensors

```{r getSensors}
  getSensors( uri="http://www.opensilex.org/demo/2018/s18001")
```

## Vectors

```{r getVectors}
  getVectors( serialNumber = "01BD1DD71500001")
```

## Events

```{r getEvents}
# Where to find the "type" URL ?
# Where is the ontology access  https://github.com/OpenSILEX/ontology-vocabularies/blob/master/oeev.owl
  getEvents( type = "http://www.opensilex.org/vocabulary/oeev#Fertilization", pageSize=10)
```

## Annotations

```{r getAnnotations, eval=FALSE}

  getAnnotations( comment = "Ustilago maydis infection" , pageSize=10)
```

# Fetch Data from the webservice

## Environment data

```{r getData}
getData(variableUri = "http://www.opensilex.org/demo/id/variables/v014")$data
```

```{r getEnvironmentData}
  getEnvironmentData( variable = "http://www.opensilex.org/demo/id/variables/v010", pageSize=50)$data
```

## Phenotypic data

```{r getPhenotypeData}
  head(getPhenotypeData( 
                        variable = "http://www.opensilex.org/demo/id/variables/v001", 
                        pageSize=10)$data)
```
 
# Send data to the web service

Here we will detail the so called POST functions, that are used to send data to the web service.

## Annotations

```{r postAnnotations}
  postAnnotations(    
    creator = "http://www.opensilex.org/demo/id/agent/marie_dupond",
    motivatedBy = "http://www.w3.org/ns/oa#commenting",
    bodyValues = list(paste("the object has been observed", rpois(n = 1, lambda = 5), sep = "")),
    targets = list("http://www.phenome-fppn.fr/test/2019/o19000074")
  )
```

## Data

Used to send data individually (usefull for testing if your dataset is correctly formated)

```{r postData}
postData(
provenanceUri = "http://www.opensilex.org/demo/2018/pv181515071552",
objectUri = "http://www.opensilex.org/demo/2018/o18000076",
variableUri = "http://www.opensilex.org/demo/id/variable/v0000001",
date = "2017-06-15T10:30:00+0200",
value = 7.3)
   
```

## Dataset

A more usefull version of postData that takes a dataset as input, with the correct column names

```{r postDataset}

attri =data.frame(   
  provenanceUri = c("http://www.phenome-fppn.fr/test/id/provenance/1569422784579",
                    "http://www.phenome-fppn.fr/test/id/provenance/1569422784579",
                    "http://www.phenome-fppn.fr/test/id/provenance/1569422784579",
                    "http://www.phenome-fppn.fr/test/id/provenance/1569422784579"),
  objectUri = c("http://www.phenome-fppn.fr/test/2019/o19000076",
                "http://www.phenome-fppn.fr/test/2019/o19000076",
                "http://www.phenome-fppn.fr/test/2019/o19000076",
                "http://www.phenome-fppn.fr/test/2019/o19000076"),
  variableUri = c("http://www.phenome-fppn.fr/test/id/variables/v027",
                  "http://www.phenome-fppn.fr/test/id/variables/v027",
                  "http://www.phenome-fppn.fr/test/id/variables/v027",
                  "http://www.phenome-fppn.fr/test/id/variables/v027"),
  date = c("2017-06-15T01:45:00+0200",
           "2017-06-15T02:46:00+0200",
           "2017-06-15T03:47:00+0200",
           "2017-06-15T04:48:00+0200"),
  value = c(11,
            22,
            33,
            44)
)
  postDataset(data = attri)
   
```

## Events	

```{r postEvents}
  postEvents(    
    rdfType = "http://www.opensilex.org/vocabulary/oeev#MoveFrom",
    concernedItemsUris = list("http://www.phenome-fppn.fr/test/2019/o19000076"),
    date = "2017-09-08T12:00:00+01:00",
    properties = list(
      rdfType = "http://xmlns.com/foaf/0.1/Agent",
      relation = "http://www.opensilex.org/vocabulary/2018#hasContact",
      value= "http://www.opensilex.org/demo/id/agent/marie_dupond"
    ),
    description = "The description of your event",
    creator = "http://www.opensilex.org/demo/id/agent/marie_dupond"
  )
```

## Provenances	

```{r postProvenances}
  postProvenances(
   label = "insertionProvenance_label",
   comment = "comment my provenance",
   metadata = list(meta1 = "15", meta2 = "owner1")
   )
```

## ScientificObjects	

Inserting Scientific objects has two variants, one accepting a list of lists, and the other beeing a data.frame with appropriate column names for the properties.
**Important** the alias block (rdfs:label) is mandatory in order to insert a new scientific object.

```{r postScienticObjects}
postScientificObjects(
  rdfType = "http://www.opensilex.org/vocabulary/oeso#Plot",
  experiment="http://www.opensilex.org/demo/DIA2017-1",
  properties=list(
    list(
      rdfType = "http://www.opensilex.org/vocabulary/oeso#Species",
      relation = "http://www.opensilex.org/vocabulary/oeso#hasSpecies",
      value = "http://www.phenome-fppn.fr/id/species/triticumaestivum"),
    list(
      rdfType = NA,
      relation ="http://www.w3.org/2000/01/rdf-schema#label",
      value ="objectAlias")
)
)
```

## Sensors	

```{r postSensors}
postSensors(rdfType = "http://www.opensilex.org/vocabulary/oeso#Spectrometer",
            label = "Sensor_label",
            brand = "Sensor_brand",model = "Sensor_model",serialNumber = "",
            inServiceDate = "2017-06-15",dateOfPurchase = "2017-06-15",
            dateOfLastCalibration = "2017-06-15",personInCharge = "admin@opensilex.org"
)
```

## Traits

```{r postTraits}
  postTraits(
   label = "insertionTrait_label",
   comment = "comment my trait",
   ontologiesReferences = list(property = "http://www.w3.org/2004/02/skos/core#closeMatch",
                   object = "http://www.cropontology.org/rdf/CO_715:0000139",
                   seeAlso = "http://www.cropontology.org/ontology/CO_715/")
                   )
```
## Methods	

```{r postMethods}
  postMethods(
    uri = "http://www.opensilex.org/demo/id/units/m016",
  label = "insertionmethod_label",
  comment = "comment my method",
  ontologiesReferences = list(
     property = "http://www.w3.org/2004/02/skos/core#closeMatch",
     object = "http://www.cropontology.org/rdf/CO_715:0000139",
     seeAlso = "http://www.cropontology.org/ontology/CO_715/")
  )
```


## Units	

```{r postUnits}
  postUnits(
   label = "insertionUnit_label",
   comment = "comment my unit",
   ontologiesReferences = list(property = "http://www.w3.org/2004/02/skos/core#closeMatch",
                   object = "http://www.cropontology.org/rdf/CO_715:0000139",
                   seeAlso = "http://www.cropontology.org/ontology/CO_715/")
                   )
                   
```

## Variables	

```{r postVariables}
postVariables(uri="http://www.opensilex.org/demo/id/variables/v0042",
              label = "insertionVariable_label",
              comment = "comment my variable",
              ontologiesReferences = list(
                property = "http://www.w3.org/2004/02/skos/core#closeMatch",
                object = "http://www.cropontology.org/rdf/CO_715:0000139",
                seeAlso = "http://www.cropontology.org/ontology/CO_715/"),
              trait = "http://www.opensilex.org/demo/id/traits/t016", 
              method = "http://www.opensilex.org/demo/id/units/m016", 
              unit = "http://www.opensilex.org/demo/id/units/u007")
  
```

## Vectors

```{r postvectors}
postVectors(
  uri="http://www.opensilex.org/demo/2019/v1909",
  rdfType = "http://www.opensilex.org/vocabulary/oeso#UAV",
  label = "Vector_label",
  brand = "Vector_brand",
  serialNumber = "1477",
  inServiceDate = "2017-06-15",
  dateOfPurchase = "2017-06-15",
  personInCharge = "admin@opensilex.org"
)
```

 
 ------------ Developper part  ------------  
 
# Get more informations on during a call
## Can be useful to understand what append ruing a call
```{r setLog level}
  setLogLevel("DEBUG")
  getEnvironmentData( variable = "http://www.opensilex.org/demo/id/variables/v050", pageSize=50)
  setLogLevel("INFO")
```
## Session info

```{r session,echo=FALSE}
  sessionInfo()
```

# References
1. R Development Core Team (2018). R: A language and environment for statistical computing. R Foundation for
      Statistical Computing, Vienna, Austria. ISBN 3-900051-07-0, URL http://www.R-project.org.

