---
title: "Requests on the web service"
author: "I.Sanchez & A. Charleroy (INRA MISTEA)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Requests on the web service}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!---
Program  : requestWS.Rmd
Author   : I.Sanchez (MISTEA)
Objective: a vignette for phisWSClientR library
           an introduction to requesting the phenomeapi (phis web service)
Creation : 19/03/2018
Update   : 14/02/2019
-->

***

<div style="background-color:rgba(0, 255, 0,0.1); text-align:left; vertical-align: center; padding:10px 0;">
This tutorial explains how to access to the first **phenomeapi** web service (PHIS system information).

For this first web service, the available functions are:

* getEnvironment()
* getExperiments()
* getImageAnalysis()
* getProjects()
* getVariableByCategory()
* getWatering()
</div>


```{r,echo=TRUE,message=FALSE, warning=FALSE}
  library(phisWSClientR)
```

# Creation of an identifier
Ask permission to request to the web service:

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  # If you want to access to the public web service 
  connect(apiID="ws_public")
```

```{r,echo=TRUE,eval=FALSE,message=FALSE, warning=FALSE}
  # If you want to access to a private web service, you have to insert the address of the WS and the port
  connect(apiID="ws_private",url = "147.99.7.5:8080/phenomeapi/resources/")
```


*aToken* will be used in all the requests to the web service.

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  aToken<-getToken("guestphis@supagro.inra.fr","guestphis")
```

# Retrieves the available projects
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getProjects(aToken$data)$data
  
  getExperiments(aToken$data,projectName = "XYZ")
```

The user can define a variable with the experiment URI provided by the result of *getExperiments()* function that can be used in the call of each functions useful to retrieve the data:
  
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  myExp<-getExperiments(aToken$data,projectName = "XYZ")$data$experimentURI
  print(myExp)
```

# Retrieves the variables of an experiment by category
Different categories of variables are available in the phenomeapi web service:

<div style="background-color:rgba(0, 255, 0,0.1); text-align:left; vertical-align: center; padding:10px 0;">
* environment
* imagery
* watering
* weighing
</div>

## Environment
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  vars <- getVariablesByCategory(aToken$data ,category = "environment",experimentURI = myExp)

  # Only the first rows of the data.frame containing the 'environment' variables...
  head(vars$data)
```

## Imagery
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  # first rows of 'imagery' variables
  head(getVariablesByCategory(aToken$data,category ="imagery",experimentURI=myExp)$data)
```

## Watering
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  # first rows of 'watering' variables
  head(getVariablesByCategory(aToken$data,category ="watering",experimentURI=myExp)$data)
```


# Retrieves the imagery data of an experiment
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  # first we count the number of observation in the dataset
  tpCount<-getImagesAnalysis(token=aToken$data, experimentURI=myExp,
                             variablesName = list("plantHeight"),
                             labelView ="side0",
                             verbose=FALSE)$totalCount
  # we retrieve this number with $totalcount
  # In this example, we set the pageSize to 10 for speed purpose but in real example, please use tpCount
  requetManip<-getImagesAnalysis(token=aToken$data, experimentURI=myExp,
                             variablesName = list("plantHeight"),
                             labelView ="side0",
                             pageSize=10)$data
  head(requetManip)

  # getPlants function allows the user to retrieve all the information for all the plants of an experiment
  # In this example, we set the pageSize to 10 for speed purpose but in real example, please use tpCount
  idManip<-getPlants(token=aToken$data, experimentURI=myExp,
                     pageSize=10)$data
  head(idManip)
```

# Retrieves the environment data of an experiment
try the following code and explore the dataset

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  myFacility<-"http://www.phenome-fppn.fr/m3p/es2"
  tpCount<-getEnvironment(token = aToken$data ,
                     experimentURI = myExp,
                     facility=myFacility,
                     variables = "air humidity_weather station_percentage")$totalCount
  #--- Air humidity
  # we retrieve this number with $totalcount
  # In this example, we set the pageSize to 10 for speed purpose but in real example, please use tpCount
  myMeteoHumidity<-getEnvironment(token = aToken$data,experimentURI=myExp,
                     facility=myFacility,
                     variables="air humidity_weather station_percentage",
                     pageSize=10)$data
```

# Retrieves the watering data of an experiment
try the following code and explore the dataset

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  tpCount<-getWatering(token=aToken$data,experimentURI=myExp,
                       variablesName = list("weightAfter"),
                       verbose=FALSE)$totalCount
  # we retrieve this number with $totalcount
  # In this example, we set the pageSize to 10 for speed purpose but in real example, please use tpCount
  requetIrrig<-getWatering(token=aToken$data,experimentURI=myExp,
                           variablesName = list("weightAfter"),
                           pageSize=10)$data
```

# Session info

```{r session,echo=FALSE,message=FALSE, warning=FALSE}
  sessionInfo()
```

# References
1. R Development Core Team (2015). R: A language and environment for statistical computing. R Foundation for
      Statistical Computing, Vienna, Austria. ISBN 3-900051-07-0, URL http://www.R-project.org.
