---
title: "Requests on the web service 2 - OpenSILEX"
author: "I.Sanchez, A. Charleroy, Jean-Eudes Hollebecq (INRA MISTEA)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Requests on the web service 2 - OpenSILEX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!---
Program  : introWS2.Rmd
Author   : I.Sanchez (MISTEA)
Objective: a vignette for phisWSClientR library
           an introduction to requesting the phenomeapi (phis web service) WS2
Creation : 24/01/2019
Update   : 14/02/2019
-->

***

<div style="background-color:rgba(0, 255, 0,0.1); text-align:left; vertical-align: center; padding:10px 0;">
This tutorial explains how to access to the second **phenomeapi** web service (PHIS system information).

For this second web service, the available functions are:

* getAnnotations()
* getEnvironmentData()
* getEvents()
* getExperiments2()
* getInfrastructures()
* getMethods2()
* getPhenotypeData()
* getProjects2()
* getScientificObjects()
* getSensors()
* getSpecies()
* getTraits()
* getUnits()
* getVariables2()
* getVectors()
</div>


```{r,echo=TRUE,message=FALSE, warning=FALSE}
  library(phisWSClientR)
```

# Creation of an identifier

Ask permission to request to the web service:

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  # If you want to access to a private web service, you have to insert the address of the WS and the port
  connectToPHISWS(apiID="ws_private",username="guest@opensilex.org",password="guest", url = "www.opensilex.org/openSilexAPI/rest/")
```

*aToken* is your identification and grants the access to some data according to your login information.

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  aToken <- opensilexWSClientR::getToken("guest@opensilex.org","guest")
```

# Fetch informations about elements in the web service

## Projects

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getProjects2( uri="http://www.opensilex.org/demo/PHENOME-FPPN")
```

## Experiments

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getExperiments2( uri="http://www.opensilex.org/demo/DIA2017-1")
```
## Scientific Object

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getScientificObjects( uri="http://www.opensilex.org/demo/2018/o18000076")
```

##Infrastructures

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getInfrastructures(uri="https://emphasis.plant-phenotyping.eu")
```

## Variables 

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getVariables2( uri="http://www.opensilex.org/demo/id/variables/v010")
```

## Traits

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getTraits( uri="http://www.opensilex.org/demo/id/traits/t010")
```

## Methods

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getMethods2( uri="http://www.opensilex.org/demo/id/methods/m010")
```

## Units

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getUnits( uri="http://www.opensilex.org/demo/id/units/u007")
```

## Species

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getSpecies( uri="http://www.opensilex.org/id/species/helianthusannuus", language = "en")
```

## Sensors

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getSensors( uri="http://www.opensilex.org/demo/2018/s18001")
```

## Vectors

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getVectors( serialNumber = "01BD1DD71500001")
```

## Events

```{r,echo=TRUE,message=FALSE, warning=FALSE}
# Where to find the "type" URL ?
# Where is the ontology access  https://github.com/OpenSILEX/ontology-vocabularies/blob/master/oeev.owl
  getEvents( type = "http://www.opensilex.org/vocabulary/oeev#Fertilization", pageSize=10)
```

## Annotations

```{r,echo=TRUE,message=FALSE, warning=FALSE}
# Where to find the "type" URL ?
# Where is the ontology access https://github.com/OpenSILEX/ontology-vocabularies/blob/master/oeev.owl
  getAnnotations( comment = "Ustilago maydis infection" , pageSize=10)
```

# Fetch Data from the webservice

## Environment data

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  getEnvironmentData( variable = "http://www.opensilex.org/demo/id/variables/v010", pageSize=50)
```

## Phenotypic data

```{r,echo=TRUE,message=FALSE, warning=FALSE}
  head(getPhenotypeData( 
                        variable = "http://www.opensilex.org/demo/id/variables/v001", 
                        pageSize=10)$data)
```
 ------------ Developper part  ------------
# Get more informations on during a call
# Can be useful to understand what append ruing a call
```{r,echo=TRUE,message=FALSE, warning=FALSE}
  setLogLevel("DEBUG")
  getEnvironmentData( variable = "http://www.opensilex.org/demo/id/variables/v010", pageSize=50)
  setLogLevel("INFO")
```
# Session info

```{r session,echo=FALSE,message=FALSE, warning=FALSE}
  sessionInfo()
```

# References
1. R Development Core Team (2018). R: A language and environment for statistical computing. R Foundation for
      Statistical Computing, Vienna, Austria. ISBN 3-900051-07-0, URL http://www.R-project.org.

